Inferring true changes in microbial abundance from taxonomically-biased microbiome measurements

### 2021-11-16 abundance-measurement.Rmd
Edit abundance-measurement section

We distinguish between two types of species-level *relative abundances*---abundances in which the absolute scale is not intrinsically meaningful.
The first type consists of the *ratios* among the abundances of two or more species, such as a ratio of 5:4:2 among three species.
The second type consists of the *proportion* of an individual species, equal to its abundance divided by that of the chosen root taxon (e.g. Prokaryotes).
Species' ratios and proportions constitute distinct views of relative abundance and hence what it means for relative abundances to change across samples.
Although proportions are dominant in microbiome DA analysis, researchers increasingly adopt the ratio view via analyses adapted from the statistical framework of Compositional Data Analysis (CoDA).
The different manner in which taxonomic bias affects ratios vs. proportions (@mclaren2019cons) forms the basis in subsequent sections for understanding the extent to which its effects cancel in DA analysis.

@@

**Difference between species proportions and species ratios:**
Consistent taxonomic bias at the species level creates a consistent fold error in the ratios among species, but a varying fold error in the proportions of individual species.
But a proportion is, after all, just a ratio between two taxa in which the denominator taxon is the entire taxonomic domain under study.
The difference in behavior from that of ratios between species arises because the efficiency of higher-level taxa, which consist of the aggregate of many species, varies across samples with the changing relative abudances among its constituent species.

>>>>

We distinguish between two types of species-level *relative abundances*.
The first type consists of the *ratios* among the abundances of two or more species, such as a ratio of 5:4:2 among three species.
The second type consists of the *proportion* of an individual species, equal to its abundance divided by that of a particular root taxon (e.g. Prokaryotes).
DA analyses of proportions consider how the proportion of a species changes indepently of other species, whereas DA analyses of ratios consider the conserted change in relative abundance among two or more species.
Although proportion-based analysis is dominant, microbiome researchers increasingly use ratio-based analyses adapted from the statistical framework of Compositional Data Analysis (CoDA).
The different manner in which taxonomic bias affects ratios vs. proportions (@mclaren2019cons) forms the basis in subsequent sections for understanding the extent to which its effects cancel in DA analysis.

@@

**Difference between species proportions and species ratios:**
Consistent taxonomic bias at the species level creates a consistent fold error in the ratios among species, but a varying fold error in the proportions of individual species.
But a proportion is, after all, just a ratio between two taxa in which the denominator taxon is the entire taxonomic domain under study.
The difference in behavior from that of ratios between species arises because the efficiency of higher-level taxa, which consist of the aggregate of many species, varies across samples with the changing relative abundances among its constituent species.
